---
title: "R Notebook For John"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

#Load packages
```{r, message = FALSE}
#install.packages("randomForest")
#install.packages("prediction")
#install.packages("class")
#install.packages("rpart")
#install.packages("olsrr")
#install.packages("broom")
#install.packages("modelr")
#install.packages("lme4")
#install.packages("lmerTest")


library(randomForest)
library(prediction)
library(class)
library(rpart)
library(olsrr) #John, load this package. It is way better for linear models. no need to separately get your confidence intervals!
library(tidyverse) #Also, tidyverse is an essentail. makes manipulating and reading data a easy (takes a while to download)
library(broom) #I forget what this is for! but use it anyways!
library(modelr) #For hierarchical modeling
library(lme4) #For Hierarchical modeling
library(lmerTest) # <- gives hierarchical modeling p-values (There is a reason it's NOT with lme4)



```

#Load data
```{r, message = FALSE}
X2017_All <- read_csv("John Dataset.csv") # <-  from tidyverse package
```
#Summary stats
```{r}

#check for outliers and set filter for interquartiles for one model
summary(X2017_All)
```

#Charge_Day Histogram
```{r}
ggplot(data = X2017_All, aes(x = Charge_Day)) +
  geom_histogram(binwidth = 1)
```

#Charge Week Histogram
```{r}
ggplot(data = X2017_All, aes(x = Charge_Week)) +
  geom_histogram(binwidth = 1)

```


#Non-standardized model
We don't need to standardize this because R does this for us! Keep it regular! 
```{r}
#######################################################################
###Main Model, seems to be working best when I don't remove outliers###
#######################################################################
All<- lm(Charge_Day ~  Day_Number+ Week_Number  + Charge_Week , data = X2017_All)
ols_regress(All) # <- from olsrr
#summary(All)
#confint(All)

  
```

```{r}
#######################################################################
###Main Model, seems to be working best when I don't remove outliers###
#######################################################################
All<- lm(CDZ ~  Day_Number+ Week_Number  + CWZ , data = X2017_All)
ols_regress(All) # <- from olsrr
#summary(All)
#confint(All)

  
```

```{r}
ggplot(data = X2017_All, aes(x = CWZ, y = CDZ)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}
ggplot(data = X2017_All, aes(x = Charge_Week, y = Charge_Day)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#correlation
Proof you don't need to standardize
```{r}
cor(X2017_All$Charge_Day, X2017_All$Charge_Week)
cor(X2017_All$CDZ, X2017_All$CWZ)

```

#Let's try hierarchical modeling...

```{r}


X2017_All <- X2017_All %>% 
  arrange(Week_Number, Day_Number) %>% 
  mutate(week.f = factor(Week_Number),
         day.f = factor(Day_Number), 
         provider.f = factor(Provider_Name)) %>% 
  arrange(Provider_Name, Week_Number, Day_Number)
```

#Fixed effects only
```{r}
mod1 = lmer(Charge_Day ~ 1 + Charge_Week + (1|provider.f), REML = TRUE, data = X2017_All)
summary(mod1)
```

#include week & day as fixed effects
This is a large dataset... it might take a while
```{r}
mod2 = lmer(Charge_Day ~ 1 + Charge_Week + Week_Number + Day_Number +(1|provider.f), REML = TRUE, data = X2017_All)
summary(mod2)
```




#Problem with this model....
```{r}
mod3 = lmer(Charge_Day ~ 1 + Charge_Week + Week_Number + Day_Number +(1+ Charge_Week |provider.f), REML = TRUE, data = X2017_All)
summary(mod3)
```